---
title: "p8105_hw6_rl3328"
author: "Ruixi Li"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r load_libraries}
library(modelr)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}


homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)


model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Problem 2

## load data
```{r q2_data_cleaning}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())


```

## create 5000 bootstrap samples

```{r bootstrap}

boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

weather_boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )

boot_straps = 
  tibble(strap_number = 1:5) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )

```

## quick check
```{r quick_check, include=FALSE}
weather_boot_straps |> 
  slice(1:3) |> 
  mutate(strap_sample = map(strap_sample, arrange, tmin)) |> 
  pull(strap_sample)
```

## get estimates

```{r estimates}
bootstrap_results = 
  weather_boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    tidy = map(models, broom::tidy),
    glance = map(models, broom::glance))|> 
  select(-strap_sample, -models) |> 
  unnest(c(tidy,glance), names_repair = "universal") |> 
  select(strap_number,term,r.squared,estimate) |>
  filter(term !="(Intercept)") |>
  group_by(strap_number) |>
        pivot_wider(
    names_from = term,
    values_from = estimate) |>
  mutate(log = log(tmin*prcp)) |>
  select(-tmin,-prcp) |>
  janitor::clean_names()
  



a =
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    tidy = map(models, broom::tidy),
    glance = map(models, broom::glance))|> 
  select(-strap_sample, -models) |> 
  unnest(c(tidy,glance), names_repair = "universal") |> 
  select(strap_number,term,r.squared,estimate) |>
  filter(term !="(Intercept)") |>
  group_by(strap_number) |>
        pivot_wider(
    names_from = term,
    values_from = estimate)
```

## plot the distribution

```{r plot}

r2_dist = bootstrap_results |>
  ggplot() +
  geom_histogram(aes(x = r_squared), color = "blue") 
r2_dist
ggsave("r2_dist.png")

log_dist = bootstrap_results |>
  ggplot() +
  geom_histogram(aes(x = log), color = "red") 
log_dist
ggsave("log_dist.png")
```

Description for the plots:

* r2_dist: the r2 for the 5000 bootstrap samples are approximately normally distributed.The mean or mode of r2 falls around 9.1 and the variance is not large(most of data locates between 0.87 and 0.95). These suggests that a high proportion of variance in the `tmax`  can be explained by `tmin` and `prcp` included in the model.The relatively tight clustering of r2 values also suggests that the underlying data are likely to be of high quality with strong linear relationships.


* log_dist: through log transformation, 3288 data (beta1*beta2<0) was excluded; meanwhile, because beta1Xbeta2 is too small(positive), log transformation reduced skewness and made the distribution of this product more symmetrical. But the overall distribution is still left skewed. most of the data falls around -6.These indicates that for more than 3288 samples, `tmin` and `prcp` have opposite effects on `tmax`; for 1822 samples, `tmin` and `prcp` have effects on `tmax`in the same direction. Absolute of beta2 tend to be very small. 



## identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval 

```{r}
# r2

# get quantiles
quantiles = round(quantile(pull(bootstrap_results,r_squared), probs = c(0.025, 0.975)),2)

# identify the 2.5% and 97.5% quantiles
lower_bound <- quantiles[1]  # 2.5% quantile
upper_bound <- quantiles[2]  # 97.5% quantile

# paste them to create the 95% Confidence Interval
cat("95% Confidence Interval for r2: [", lower_bound, ", ", upper_bound, "]", sep="")

# log(b1*b2)

# get quantiles
quantiles2 = round(quantile(pull(bootstrap_results,log), probs = c(0.025, 0.975), na.rm= TRUE),2)

# identify the 2.5% and 97.5% quantiles
lower_bound2 <- quantiles[1]  # 2.5% quantile
upper_bound2 <- quantiles[2]  # 97.5% quantile

# paste them to create the 95% Confidence Interval
cat("95% Confidence Interval for r2: [", lower_bound2, ", ", upper_bound2, "]", sep="")
```

# Problem 3

```{r load_libs}
library(stats)
library(car)
```

## Data cleaning

```{r clean_data}
bw = read_csv("data/birthweight.csv")
sum(is.na(bw)) # there's no missing value

# Convert some numeric variables to factors
bw <- bw |>
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("male", "female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    malform = factor(malform, levels = c(0, 1), labels = c("absent", "present")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other"))
  )

```

## build up the regression model using stepwise method

```{r}
full_model <- lm(bwt ~ ., data = bw)
stepwise_model <- step(full_model, direction = "both")
summary(stepwise_model)

```



## model diagnostics1

```{r}
# Add predictions and residuals to the data
bw_diagnostic1 <- bw |>
  add_predictions(stepwise_model) |>
  add_residuals(stepwise_model)

# Plot model residuals against fitted values
bw_diagnostic1 |> ggplot( aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs. Fitted Values"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")

```
* I first fit an initial full regression model including all potential predictors listed. Then I applied a stepwise regression approach with both direction to select a set of variables that contribute meaningfully to the model according to AIC. I drew a residual vs. fit plot, which indicates that the stepwise model's residuals spread randomly around the horizontal line at 0 (the red dashed line) and have no pattern. But I detected assumption violation of heteroscedasticity and some potential influential outliers. 
So, I tried to transform the bwt into log/square root(bwt) to see if this fix the heteroscedasticity.

## transform bwt into log format and repeat the modeling process 
```{r}
# remember to rerun previous code chunk `clean_data` to get the original bw dataframe

log_full_model <- lm(log(bwt) ~ ., data = bw)
log_stepwise_model <- step(log_full_model, direction = "both")
summary(log_stepwise_model)

```

## model diagnostics2

```{r}
# Add predictions and residuals to the data
bw_diagnostic2 <- bw |>
  add_predictions(log_stepwise_model) |>
  add_residuals(log_stepwise_model)

# Plot model residuals against fitted values
bw_diagnostic2 |> ggplot( aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs. Fitted Values"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")

```




# check multicollinearity
```{r}
vif_results <- vif(log_stepwise_model) 
print(vif_results)

```
Acorrding to the results of VIF analysis and my knowlege, there's relationship between bhead and blength,delwt and ppwt. I would like to keep bhead and ppwt in my model and delete the other two covariates.

## modify the model based on my knowledge

```{r}
my_model = update(log_stepwise_model, . ~ . - blength - delwt)
summary(my_model)

```

## model diagnostics 3
```{r}
# remember to rerun previous code chunk `clean_data` to get the original bw dataframe

# Add predictions and residuals to the data
bw_diagnostic3 <- bw |>
  add_predictions(my_model) |>
  add_residuals(my_model)

# Plot model residuals against fitted values
bw_diagnostic3 |> ggplot( aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs. Fitted Values"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")

```

* According to the plot, this model seems good with no violation of assumption, but still have some outliers/influential points.

## identify potential outliers or influential points

```{r}
influencePlot(my_model)
```
* According to studentized residual(absolute StudRes > 2), point 1477,1748 and 2691 are outliers.
* According to leverage/Hat value(>3*(p/n)=0.0083),point 1,1477,1993,2691 are outliers.
* According to Cook's distance(>1),

* Outliers: Observations 1477, 1748, and 2691 are likely outliers.
* Influential Points: Observations 1477 and 1748 are likely influential due to their high Cook's distances. Observation 1993 could also be influential due to its very high leverage, even though it's not an outlier by the residual criterion.


## remove point 1477 and 1748 from the observations
```{r}
# remember to rerun previous code chunk `clean_data` to get the original bw dataframe
bw = bw[-c(1477,1748),]

log_full_model <- lm(log(bwt) ~ ., data = bw)
log_stepwise_model <- step(log_full_model, direction = "both")
log_mymodel = update(log_stepwise_model, . ~ . - blength - delwt)
summary(log_mymodel)
```

* After removing the influential points, the coefficient of fincome experienced significant change, and the direction altered(from negative value to positive value), which makes more sense igven common sense.

## model diagnostics 4
```{r}
# remember to rerun previous code chunk `clean_data` to get the original bw dataframe

# Add predictions and residuals to the data
bw_diagnostic4 <- bw |>
  add_predictions(log_mymodel) |>
  add_residuals(log_mymodel)

# Plot model residuals against fitted values
bw_diagnostic4 |> ggplot( aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Residuals vs. Fitted Values"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red")


influencePlot(log_mymodel)
```

* Although there's still influential outliers, I won't remove them any longer.(removing point 2689 didn't change the model so much)

## model comparison

### model 1
```{r}
model_length_gestage <- lm(bwt ~ blength + gaweeks, data = bw)


model_head_length_sex_interactions <- lm(bwt ~ bhead * blength * babysex, data = bw)
```

### model 2